### 2016-09-05 What Does Reproducible Research Mean for Plant Pathology.Rmd
Remove redundant files, add links in README.Rmd, fix links in manuscript

---
title: "What Does Reproducible Research Mean for Plant Pathology?"
author: "A. H. Sparks, E. M. Del Ponte, N. J. Grünwald, and Z. Foster"
date: "`r Sys.Date()`"
output:
  word_document:
    reference_docx: Phytopathology_Style_Reference.docx
bibliography: bibliography.bib
csl: phytopathology.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. Adam H. Sparks: Centre for Crop Health, University of Southern Queensland, 
Toowoomba, Qld 4350, Australia
2. Emerson M. Del Ponte: Universidade Federal de Viçosa, Viçosa, MG, Brasil
3. Niklaus J. Grünwald: Horticultural Crops Research Laboratory, USDA
Agricultural Research Service, Corvallis, OR 97330, USA
4. Zachary Foster: Horticultural Crops Research Laboratory, USDA Agricultural
Research Service, Corvallis, OR 97330, USA

# ABSTRACT

Abstracts are mandatory and limited to one 200 word paragraph.

# MAIN TEXT
Reproductibility and replicability in scientific research have once again been
highlighted recently [@Nature_Editorial_2016; @Baker2016a] as an issue.

Patil et al. [-@Patil066803] have provided several definitions to clarify the
concepts surrounding reproducibility and replicability. For the purposes of this
paper we follow the definitions as given by Patil et al. [-@Patil066803].

* Why reproducible research

## BEST METHODS FOR REPRODUCIBLE RESEARCH
* Provide definitions (provide defintions for terms used so it's clear)
* Data
  * Data formatting (flat files; use Comma Chameleon, Table Tool, others?)
  * Data storage (don't edit raw data files; use file permissions to prevent changes to raw data files, use data bases where possible and appropriate; etc.)
* When publishing
  * Provide data
  * Provide code
* Using GitHub for code (and small data?)
* Using Figshare or Zenodo vs a lab website (DOIs, other reasons)

## WHAT IS THE STATE OF REPRODUCIBLE RESEARCH IN PLANT PATHOLOGY?
* Madden et al. [-@Madden2015] supply an *e-**X**tra*\* with repoducible examples
for readers.
* Duku et al. [-@Duku2016] provide models, data and code, (http://adamhsparks.github.io/MICCORDEA/) necessary to
replicate the entire study modelling the effects of climate change on rice
bacterial blight and rice leaf blast in Tanzania.
* Sparks et al. [-@Sparks2011; -@Sparks2014] provide models, data and code, (http://adamhsparks.github.io/Global-Late-Blight-MetaModelling/)
necessary to replicate model development and the subsequent the study on the effects
of climate change on potato late blight.
* Other examples from plant pathology providing e-Xtras or supplemental material

## DISCUSSION

## ACKOWLEDGEMENTS

### Notes
https://www.r-statistics.com/2016/07/the-reproducibility-crisis-in-science-and-prospects-for-r/

### LITERATURE CITED