### 2017-04-04 reproducibility_criteria.Rmd
new; adapted from issue #5

---
title: "Criteria for evaluating reproducibility"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

* Raw data availability
    0. Not available or not mentioned in the publication
    1. Available upon request to author
    2. Online, but inconvenient/non-permanent (e.g. login needed, pay wall, FTP server, personal lab website)
    3. Freely available online to anonymous users for foreseeable future
* Raw data annotation
    0. No annotation
    1. The structure/content of the data is clear, but the relevance to the study is not
    2. Fully understandable in the context of the paper
    3. Fully understandable without reading the paper
* Raw data tidiness
    0. Not easily parsed with code or loaded by standard office software (e.g. unstructured raw text or binary equipment outputs or no raw data)
    1. Difficult, but possible, to parse with code or manipulate with standard office software (i.e. you technically could, but would not want to)
    2. Able to parse using code with minimal cleanup (e.g. you might need a few extra lines of code for clean up or do a few extra steps to work with your operating system. An example would be a text file with multiple tables)
    3. Tidy data in a standardized format (e.g. tsv) that is loaded correctly without cleanup or using uncommon parsing tools.
* Computational methods availability
    0. Not available or not mentioned in the publication
    1. Available upon request to author
    2. Online, but inconvenient/non-permanent (e.g. login needed, pay wall, FTP server, personal/lab website)
    3. Freely available online to anonymous users for foreseeable future
* Availability of software used
    0. Not available or not mentioned in the publication
    1. Uses expensive proprietary software that only institutions would typically purchase (i.e. ArcGIS standard is $7000)
    2. Uses proprietary software that most individuals can afford (e.g. Excel, Matlab? ($500))
    3. Uses entirely open source and free software (e.g. R).
* Citation of software used
    0. Not available or not mentioned in the publication
    1. Software mentioned by name only
    2. Attempts at citing
    3. Properly cited
* Analysis automation
    0. All steps were performed manually with no choesive document describing exactly what was done
    1. Steps were performed manually, but with a detailed record (e.g. pasted commands in a text file)
    2. A set of runnable scripts were used, but have to be executed individually or the outputs require manually modifications
    3. A makefile or master script was provided that runs the entire analysis from raw data to published results