### 2017-04-10 src/README.Rmd
Update article selection based on @emdelponte
Updated to 200 articles from the journals suggested by @emdelponte in
https://github.com/adamhsparks/Reproducible-Research-in-Plant-Pathology/
issues/6

---
title: "README"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Creating a sample of articles

Twenty journals in the discipline of plant pathology were selected by the authors as being the primary choice for most plant pathologists when publishing manuscripts.

Using the assumption that journals publish an average 10 articles per issue, for a population of ~1000 articles, using a confidence level of 95% and confidence interval of 10% we need 88 samples. In [Issue #3](https://github.com/adamhsparks/Reproducible-Research-in-Plant-Pathology/issues/3), we decided to select 200 articles to have a large sample of the population.

## R setup
Set up the R environment

```{r library, message=FALSE}
library(dplyr)

set.seed(1)
```
## Import data
The list of the top 20 plant pathology journals was retrieved from Google Scholar on December 5, 2016 and saved as a CSV file for reproducibility.

```{r journal_list}
journal_list <- tibble(
  seq(1:20),
  c("Australasian Plant Pathology",
    "Canadian Journal of Plant Pathology",
    "Crop Protection",
    "European Journal of Plant Pathology",
    "Forest Pathology",
    "Journal of General Plant Pathology",
    "Journal of Phytopathology",
    "Journal of Plant Pathology",
    "Journal of Plant Virology",
    "Molecular Plant Pathology",
    "Nematology",
    "Physiological and Molecular Plant Pathology",
    "Phytoparasitica",
    "Phytopathologia Mediterranea",
    "Phytopathology",
    "Plant Disease",
    "Plant Health Progress",
    "Plant Pathology",
    "Revista Mexicana de Fitopatología",
    "Tropical Plant Pathology"))

names(journal_list) <- c("number", "publication")
```

## Create random lists
Create a randomised list of the journals

```{r journal_sample}
journals <- tibble(sample(1:20, 200, replace = TRUE))
names(journals) <- "number"
journals <- left_join(journals, journal_list, "number")
```

## Randomly select articles
Generate a random list of years between 2012 and 2016 and a random list of start pages between 1 and 150 since some journals start numbering at 1 with every issue. Then bind the columns of the randomised list of journals with the randomised years and page start numbers. This then assumes that there is no temporal effect, i.e., the time of year an article is published does not affect whether or not it is reproducible.

```{r years_and_articles}

year <- sample(2012:2016, 200, replace = TRUE)

start_page <- sample.int(150, 200, replace = TRUE)

(journals <- cbind(journals, year, start_page)[, -1])

```

## Check the number of articles per journal

```{r, count}
journals %>%  
 group_by(publication) %>% 
    tally(sort = TRUE) 
```

## R Session Info
```{r, session_info}
sessionInfo()
```