### 2017-08-15 docs/articles/reproducibility_criteria.Rmd

---
title: "Criteria for Evaluating Reproducibility"
author: "Adam H Sparks and Emerson M Del Ponte and Zachary S L Foster and Niklaus J. Grünwald"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Criteria for Evaluating Reproducibility}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Key for data collection

This vignette documents the fields and values found in the
"article_evaluations" sheet in the "article_notes" Google Sheets
workbook,
https://docs.google.com/spreadsheets/d/19gXobV4oPZeWZiQJAPNIrmqpfGQtpapXWcSxaXRw1-M/edit#gid=1699540381

### Autogenerated values

In the reproducibility score there are two parts, one part is automatically
generated, as much as possible, using the article DOI to retrieve the reference.

The articles' DOIs are looked up manually by Adam and assigned in the
"[article_notes](https://docs.google.com/spreadsheets/d/1YrIiQ3OBNjCHMlalkMi1yy8dy4hnj5dnR4PSTj7lDR4/edit#gid=0)" sheet in the googlesheets document, `doi`.

Nik has provided the five-year impact factor, `IF_5year`.

Other items including `journal`, `year`, and `contains_page` are auto-generated by the ["Assigning Articles"](https://github.com/phytopathology/Reproducible.Plant.Pathology/blob/master/vignettes/Assigning_Articles.Rmd) vignette.

#### "journal"

Auto-generated in Assigning Articles

Name of journal

Entered as character string in title case

#### "year"

Auto-generated in Assigning Articles

Year the article was published

Entered as a 4-digit number.

#### "contains_page"

Auto-generated in Assigning Articles

Randomly selected page number which is used to select article for evaluation.
The article is selected if it contains this page number in its page range.

#### "assignee"

Auto-generated in Assigning Articles

Randomly assigned evaluator selected from Adam, Emerson, Nik or Zach and
assigned to a particular article (row)

### Manually entered metadata

### "doi"

Articles are looked up by Adam using the "article_notes" sheet in googlesheets.
The DOI is entered as a character string or left as `NA` if one is not
provided.

Entered as a character string.

### "comments"

Comments column, Adam has used this to enter comments regarding article
selection in cases where the `contains_page` number does not align with
our criteria. Can be used by evaluator for other comments on the article
when evaluating.

Entered as a character string.

### "IF_5year"

Nik has looked up and provided the five-year impact factor for each journal and
entered it in the "article_notes".

Entered as a decimal value.

### Evaluation of reproducibility

#### "open"

Whether the journal is open-access or not

Entered as `TRUE`/`FALSE`/`BOTH`

#### "repro_inst"

Does the journal have guidelines for reproducibility?

- `0` - Not mentioned

- `1` - At least a suggestion to put data online, but not details on how/where.

- `2` - Detailed instructions on where to place data

- `3` - Detailed instructions on where to place data and code

#### "art_class"

What kind of research the article describes

One or more of the following:

- `Fundamental`

- `Applied`

- `Molecular`

Entered as an R `vector` object in comma-separated list with spaces.

For example: `c("Applied"", "Molecular")` or `Fundamental`

#### "comp_mthds_avail"

Score of computational methods documentation

Are the computational methods used readily available, _e.g._, R, SAS, Python
scripts are shared?

Entered as a number between `0` and `3`

- `0` - Not available and not mentioned in publication

- `1` - Available upon request to author

- `2` - Online, but inconvenient/non-permanent (_e.g._, login necessary,
pay wall, FTP server, personal lab website)

- `3` - Freely available online to anonymous users for foreseeable future
(_e.g_, GitHub, Figshare, Dataverse)

#### "software_avail"

Score of software availability

Entered as a number between `0` and `3`

- `0` - Not available or not mentioned in the publication

- `1` - Uses expensive proprietary software that only institutions would
typically purchase (_i.e._, ArcGIS standard is $7000)

- `2` - Uses proprietary software that most individuals can afford (_e.g._,
Excel, Matlab? ($500))

- `3` - Uses entirely open source and free software (_e.g._, R, Python).

#### "software_cite"

Score of citations for software used

Entered as a number between `0` and `3`

- `0` - not mentioned

- `1` - Software mentioned by name only

- `2` - Software cited with version number

- `3` - All software components (PROCs, packages) etc. cited

#### "software_used"

Entered as an R `vector` object in comma-separated list with spaces.

For example: `c("R", "TableTool")` or `R`

#### "data_avail"

Raw data availability score

Entered as a number between `0` and `3`

- `0` - Not available or not mentioned in the publication

- `1` - Available upon request to author

- `2` - Online, but inconvenient/non-permanent (_e.g._ login needed, pay wall,
FTP server, personal lab website)

- `3` - Freely available online to anonymous users for foreseeable future

#### "reproducibility_score"

Sum of the prior scores that are scored on a `0` - `3` scale. A score of `0` is
not reproducible a score of `21` is perfectly reproducible based on the
criteria agreed upon for this article.

## Bibliography fields

The following fields are automatically filled by searching for the DOI and
retrieving a BibTex entry, where applicable. In cases where a DOI is not
present or the data are incomplete or incorrect the evaluator should enter
the necessary data for the article assigned to them. In these cases, important
fields to enter are:

### "author"

Entered as character string in R `vector` format with spaces following BibTex
style, _e.g._, c("A H Sparks", "E {Del Ponte}", "Z Foster", "N Grünwald")

### "pages"

Entered as a character string in BibTex format, numbers separated by a double
"-", _e.g._, 1--18.

### "publisher"

Name of journal publisher.

Entered as character string, _e.g._ Springer or Wiley-Blackwell or
Scientific Societies

### "month"

Month in which the article was published

Entered as character string, _e.g._ jan, feb, mar...

### "number"

Issue number of the journal

Entered as an integer

### "title"

Article title

Entered as a character string

### "volume"

Volume number of the journal

Entered as an integer

## The following fields are automatically created by BibTex and are _optional_

### "category"

### "bibtexkey"

### "address"

### "annote"

### "booktitle"

### "chapter"

### "crossref"

### "edition"

### "editor"

### "howpublished"

### "institution"

### "key"

### "note"

### "organizations"

### "school"

### "series"

### "type"

### "url"

### "x.article.pm..title"