### 2021-06-02 README.rmd
Clarify directory structure in the README.md and update affiliation

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Reproducibility in Plant Pathology
![Publish Docker](https://github.com/openplantpathology/Reproducibility_in_Plant_Pathology/workflows/Publish%20Docker/badge.svg)
[![DOI](https://zenodo.org/badge/62676177.svg)](https://zenodo.org/badge/latestdoi/62676177)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)

This repository contains the data and code for our paper:

> Sparks, A.H., Del Ponte, E.M., dos Santos Alves, K., Foster, Z., Grünwald, N. (YYYY). _Title of paper_. Name of journal/book <https://doi.org/xxx/xxx>

Our pre-print is online here:

> Authors, (YYYY). _Title of paper_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to cite

Please cite this compendium as:

> Sparks, A.H., Del Ponte, E.M., dos Santos Alves, K., Foster, Z., Grünwald, N. (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for 'Status and Best Practices for Reproducible Research In Plant Pathology'_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/10.5281/zenodo.1250665>

### How to download or install

#### The R package

This repository is organized as an R package. There are custom R functions, 
`table_of_journals()` and `workflow_dia()` that are used in this repository, along with a bibliography file of the articles that were examined and the notes that are located in `inst/extdata` directory.
We have used the R package structure to help manage dependencies, to take advantage of continuous integration for automated code testing and for file organisation.

You can download the compendium as a zip from from this URL: <https://github.com/openplantpathology/Reproducibility_in_Plant_Pathology/archive/main.zip>

Or you can install this compendium as an R package, Reproducibility.in.Plant.Pathology, from GitHub with:
```{r gh-installation, eval = FALSE}
if (!require("remotes"))
  install.packages("remotes")
remotes::install_github("openplantpathology/Reproducibility_in_Plant_Pathology"
)
```

Once the download is complete, open the `Reproducibility_in_Plant_Pathology.Rproj` in RStudio to begin working with the package and compendium files.

#### The Docker Instance

Get the latest instance from Dockerhub, launch it and go to `localhost:8787` in your browser.
Login with `rstudio`, password is `rstudio`.

```
docker pull adamhsparks/reproducibility_in_plant_pathology
docker run -d -p 8787:8787
```

#### The Paper

The file structure follows a normal R package with one exception.
The top-level "/analysis" directory contains the directories and files necessary to reknit the MS Word document of the paper from an Rmd file, "/analysis/paper/paper.Rmd".

A script, `knit_paper.R` is located in `analysis/paper/knit_paper.R` that will knit the Word document

## Meta

Code: [MIT](http://opensource.org/licenses/MIT) year: `r format(Sys.Date(), "%Y")`, copyright holder: Adam H Sparks

Data: [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

Adam H Sparks
Senior Research Scientist
Farming Systems Innovation
Primary Industries Development
Department of Primary Industries and Regional Development
Level 5, 1 Nash St Perth WA 6000

<https://adamhsparks.com/>