### 2022-04-25 README.rmd
Reknit README after updating text to reflect package structure

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Reproducibility in Plant Pathology
<!-- badges: start -->
![Publish Docker](https://github.com/openplantpathology/Reproducibility_in_Plant_Pathology/workflows/Publish%20Docker/badge.svg)
[![DOI](https://zenodo.org/badge/62676177.svg)](https://zenodo.org/badge/latestdoi/62676177)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
<!-- badges: end -->

This repository contains the data and code for our article:

> Sparks, A.H., Del Ponte, E.M., Alves, K. S., Foster, Z., Grünwald, N. J. (YYYY). _Reproducibility in plant pathology: where do we stand and a way forward_. Name of journal/book <https://doi.org/xxx/xxx>

Our pre-print is online on the agriRxiv preprint server:

> Sparks, A.H., Del Ponte, E.M., Alves, K. S., Foster, Z., Grünwald, N. J. (2021). _Reproducibility in plant pathology: where do we stand and a way forward_. agriRxiv, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/10.31220/agriRxiv.2021.00082>

The paper is a systematic and quantitative review of articles published in 20 plant pathology journals that spans five years of publications.
It provides a basis for identifying what has been done so far in the discipline of plant pathology's published research to ensure computational reproducibility.
The results show that as a discipline, plant pathologists are not widely sharing data or code openly, making the works largely unreproducible.
Based on these results and our own experiences, we offer suggestions as to how we can further improve reproducibility in the discipline of plant pathology, but which are not unique to the discipline, that would allow reviewers to make better suggestions, readers to learn more from the work and earns author more citations for their work.

### How to cite

Please cite this compendium as:

> Sparks, A.H., Del Ponte, E.M., Alves, K. S., Foster, Z., Grünwald, N. J. (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for 'Status and Best Practices for Reproducible Research In Plant Pathology'_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/10.5281/zenodo.1250665>

### How to download or install

#### The R package

This repository is organized as an R package.
There is one R function, `import_notes()` that is used in the paper's figure and table making when the file, `analysis/paper/paper.Rmd` is knit.
Additionally, a bibliography file, "references.bib", of the articles that were examined and the notes from the evaluation, "Reproducibility_in_plant_pathology_notes.ods" of the articles are both located in `inst/extdata` directory.
We have used the R package structure to help manage dependencies, to take advantage of continuous integration for automated code testing and for file organisation.

You can download the compendium as a zip from from this URL: <https://github.com/openplantpathology/Reproducibility_in_Plant_Pathology/archive/main.zip>

Or you can install this compendium as an R package, Reproducibility.in.Plant.Pathology, from GitHub with:
```{r gh-installation, eval = FALSE}
if (!require("remotes"))
  install.packages("remotes")
remotes::install_github("openplantpathology/Reproducibility_in_Plant_Pathology"
)
```

Once the download is complete, open the `Reproducibility_in_Plant_Pathology.Rproj` in RStudio to begin working with the package and compendium files.

#### The Docker Instance

Get the latest instance from Dockerhub, launch it and go to `localhost:8787` in your browser.
Login with `rstudio`, password is `rstudio`.

```bash
docker pull adamhsparks/reproducibility_in_plant_pathology
docker run -d -p 8787:8787 adamhsparks/reproducibility_in_plant_pathology
```

#### The Paper

The file structure follows a normal R package with one exception.
The top-level "/analysis" directory contains the directories and files necessary to re-knit the MS Word document of the paper from an Rmd file, "/analysis/paper/paper.Rmd".

A script, `knit_paper.R`, is located in `analysis/paper/knit_paper.R` that will knit the [Word document](analysis/paper/paper.docx) in a Docker session.

## Meta

### Licensing

Code: [MIT](http://opensource.org/licenses/MIT) year: `r format(Sys.Date(), "%Y")`, copyright holder: Adam H. Sparks

Data: [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

Adam H. Sparks
Senior Research Scientist
Farming Systems Innovation
Primary Industries Development
Department of Primary Industries and Regional Development
Level 6.35, 1 Nash St., Perth WA 6000

<https://adamhsparks.netlify.app>

### Code of Conduct

Please note that the Reproducibility.in.Plant.Pathology project is released with a [Contributor Code of Conduct](https://openplantpathology.github.io/Reproducible.Plant.Pathology/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.